---
title: "Fars Functions Vignette"
author: "Swetha Garimalla"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This package is the fars function package. It is a part of Coursera's Building Packages course online. It
can be used to analyze the [US National Highway Traffic Safety Administration's Fatality Analysis Reporting System](https://www.nhtsa.gov/Data/Fatality-Analysis-Reporting-System-(FARS)) (a nationwide census of annual fatal motor vehicle accidents).

## Backround
The data used for this is primarily from the fars_data.zip folder which decompresses to accident reports in 
`*.csv.bz2` format. 

## Load Data 

In order to load the data for a single year, two functions must be used. 
`make_filename()`
`fars_read()`

```{r, echo=FALSE, results='asis'}
fn <- make_filename(2013); 
```

This assigns the value `accident_[argument].csv.bz2`, in this case `accident_2013.csv.bz2`

```{r, echo=FALSE, results='asis'}
fars_2013 <- fars_read(fn); 
```

This command reads in the single FARS file; in this case, accident_2013.csv.bz2

## Retrieves Month & Year from the Accident Year Files

### Retrieve for a Single Year

To load and explore the fatality data for a single year is, the `fars_read_years()` function must be used. 
This function can take a single year as input (shown below) and returns a list containing one element, a `tidyverse` dataframe (`tbl_df`) with the year and month of all the accidents within the year specified.  


```{r, echo=FALSE, results='asis'}
fars_2013 <- fars_read_years(2013); 
```

### Retrieve for Multiple Years

To load and explore the fatality data for multiple years, we still use the `fars_read_years()` function; however, in this case, the input is a vector of years (shown below). This function, once executed, willit return a list containing one element for each year in the input vector. Each element will be in the format of a `tidyverse` dataframe (`tbl_df`) with the year and month of all the accidents within the year specified.  


```{r, echo=FALSE, results='asis'}
fars_2014 <- fars_read_years(2013:2015); 
```

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
