---
title: "Fars Functions Vignette"
author: "Swetha Garimalla"
date: "June 10, 2018"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This package is the fars function package. It is a part of Coursera's Building Packages course online. It
can be used to analyze the [US National Highway Traffic Safety Administration's Fatality Analysis Reporting System](https://www.nhtsa.gov/Data/Fatality-Analysis-Reporting-System-(FARS)) (a nationwide census of annual fatal motor vehicle accidents).

## Backround
The data used for this is primarily from the fars_data.zip folder which decompresses to accident reports in 
`*.csv.bz2` format. 

## Load Data 

In order to load the data for a single year, two functions must be used. 
`make_filename()`
`fars_read()`

```{r, echo=FALSE, eval=FALSE}
fn <- make_filename(2013) 
fn
```

This assigns the value `accident_[argument].csv.bz2`, in this case `accident_2013.csv.bz2`

```{r, echo=FALSE, eval=FALSE}
fars_2013 <- fars_read(fn)
fars_2013
```

This command reads in the single FARS file; in this case, accident_2013.csv.bz2

## Retrieves Month & Year from the Accident Year Files

### Retrieve for a Single Year

To load and explore the fatality data for a single year is, the `fars_read_years()` function must be used. 
This function can take a single year as input (shown below) and returns a list containing one element, a `tidyverse` dataframe (`tbl_df`) with the year and month of all the accidents within the year specified.  


```{r, echo=FALSE, eval=FALSE}
fars_2013 <- fars_read_years(2013)
fars_2013
```

### Retrieve for Multiple Years

To load and explore the fatality data for multiple years, we still use the `fars_read_years()` function; however, in this case, the input is a vector of years (shown below). This function, once executed, willit return a list containing one element for each year in the input vector. Each element will be in the format of a `tidyverse` dataframe (`tbl_df`) with the year and month of all the accidents within the year specified.  


```{r, echo=FALSE, eval=FALSE}
fars_2013_15 <- fars_read_years(2013:2015)
fars_2013_15
```

## Summarizing FARS Data 
Takes specific years (either a single value denoting a single year or a vector of values denoting multiple years, like the input for the `fars_read_years()` function). The output of this function is a summary of counts of accidents for each month and year. 

```{r, echo=FALSE, eval=FALSE}
fars_summ_2013 <- fars_summarize_years(2013)
fars_summ_2013

fars_summ_2013_15 <- fars_summarize_years(2013:2015)
fars_summ_2013_15
```

## Mapping Accidents onto States

Takes input of a `state.num` (a numerical integer denoting the US state as is shown in the data set) and `year` (numerical integer denoting the year) as arguments and plots the accidents onto a map of the states. If the state number is invalid, an error is thrown. If there are no accidents in that state, a message is returned that there are no accidents to plot. Otherwise, it returns a plot of states with the number of accidents on each states the accidents based on the `year`. Returns an error if `state.num` or if 
`year` do not exist in the data set.

```{r, echo=FALSE, eval=FALSE}
library(mapdata)
fars_summ <- fars_map_state(20, 2013)
```
