---
title: "Fars Functions Vignette"
author: "Swetha Garimalla"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This package is the fars function package. It is a part of Coursera's Building Packages course online. It
can be used to analyze the [US National Highway Traffic Safety Administration's Fatality Analysis Reporting System](https://www.nhtsa.gov/Data/Fatality-Analysis-Reporting-System-(FARS)) (a nationwide census of annual fatal motor vehicle accidents).

## Backround
The data used for this is primarily from the fars_data.zip folder which decompresses to accident reports in 
`*.csv.bz2` format. 

## Load Data 

In order to load the data for a single year, two functions must be used. 
`make_filename()`
`fars_read()`

```{r, echo=FALSE, results='asis'}
fn <- make_filename(2013); #this assigns the value accident_[argument].csv.bz2, in this case     
                           #accident_2013.csv.bz2

fars_2013 <- fars_read(fn); #reads in the single FARS file; in this case, accident_2013.csv.bz2
```


## Retrieves Month & Year from the Accident Year Files

# Retrieve for a Single Year

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

# Retrieve for Multiple Years

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
